\large \bf Predicting Query Execution Time

### 2015-10-21 draft.tex
Added experiments

\usepackage{multicol}
\usepackage{graphicx}

@@

	\section{Overview of Proposed Approach}
	\includegraphics[scale=0.3]{training.png}
TODO :  
1. One more flowchart for Testing. 
2. Describe the concept of multiple models for a single operator. [MSR]
3. How do we select the best one among them. (Mention that current experiments dont use this idea, only one a simple non-linear model is used for all models with 2 features)

	Elaborate what is the plan level approach. give  a intro model
	 based techniques. i.e.,  OFF LINE TRAINING, emphasize one time.
	1. How to obtain training data (explain analyze)
	2. getting execution times at plan nodes
	3. Individual model training 
	4. Explain while testing and predict using Model more in next section

	Mention assumptions, 
	Pure in-memory.. lets solve this first.
	For example, a sort merge

join may require sorting its inputs. The sort operators are then responsible for
handing over their result to the merge join via main memory. This means that
the merge join may not require any I/O if the merge can be done purely in
main memory

@@

%%%%%%%%%%%%%  This can be followed by several other sections
	\section{Model selection and Training}
	What model is considered good ? choice of model and kernel
	If possible prove linear systems are not sufficient.
	Training data ,specification. Does more training data helps?
	 training time etc.. 
	Essentially all operators need to be covered. generation QGEN tool. 
	\section{Preliminary Experiments}

@@

	Give a background about Static and dynamic workload.  
	do experiments at 2 stages
	1. Act vs Act
	2. Est vs Est(Bias towards estimated cardinalities)
	For each type of above scenario, do testing for tpc-h queries 
	(tpc-ds in future)
	At different scales to see the model's ability to predict.
	For midterm you should be having result for 1 as well as 10 GB.
	For each above dataset, 
	Results should convey L1 (MRE) as well as queries that are within 
	[0-0.5] [0.5-1] [1-1.5] 
	\section{Conclusions and Future Work}
	Talk about generalization. 
	operator specific features
	capturing query interaction i.e., pipeline source of over-estimation
	Optimizers provides sufficient information to detect a pipeline,
	 can we use it to bind the operators and predict execution time 
	 for a set of operators instead of a single operator.

	\begin{thebibliography}{9}
	\bibitem{opthet} 
	Du, Weimin, Ravi Krishnamurthy, and Ming-Chien Shan. "Query optimization in a heterogeneous dbms." 		VLDB. Vol. 92. 1992.
	\bibitem{activeSLA}
	Xiong, Pengcheng, et al. "ActiveSLA: a profit-oriented admission control framework for database-as-a-		service providers." Proceedings of the 2nd ACM Symposium on Cloud Computing. ACM, 2011.	
	\bibitem{progress}
	Mishra, Chaitanya, and Nick Koudas. "The design of a query monitoring system." ACM Transactions on 		Database Systems (TODS) 34.1 (2009): 1.

	\bibitem{ganapathi} 
	A. Ganapathi, H. A. Kuno, U. Dayal, J. L. Wiener, A. Fox, M. I. Jordan,
	and D. A. Patterson. Predicting multiple metrics for queries: Better
	decisions enabled by machine learning. In ICDE, 2009.

@@

@@

>>>>

\usepackage{multicol}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{framed}

@@

	\section{Overview of Proposed Approach}

	To-Do : improvise diagrams with creately.com \\
	\includegraphics[scale=0.3]{training.png}
	\includegraphics[scale=0.3]{testing.png}

	In this section, we elaborate on the overview of the approach which consists of 2 phases. Like 
	most other machine learning approaches, the first phase is an off-line training phase and the 
	second is on-line testing phase.
	During the training phase, we construct number of different 
	models for each type of physical database operator(e.g.,  Sort-Merge Join, Nested Loop, Index Nested 		Loop,..). Each operator will be associated with so called 'Best-Fit' model which will be determined 
	based on the Operator model selection described in section 3.1. This model will be invoked for the
	purpose of estimating the execution time for a certain operator. Since there are only finite number of 	operators and only a couple of physical implementations for them, the space consumption is not 				excessive. 
	\\

	During the testing phase, we actually run the query to get the true cardinality estimates. While this
	seems to contradict the whole estimation problem, right now this is required because the focus is on
	solving the modeling problem given the actual cardinalities. The more practical scenario i.e., where
	we we only the know the estimated cardinalities, is a much harder problem as now the models also need 
	to be Robust to the errors in cardinality estimates. This is the natural follow-up work to do. Once 
	after the execution of the query, we need to extract features depending on the operator and invoke 
	the Best Fit model (which is already computed during training phase) to get an estimate. We sum up 
	costs for all the operator present in the plan tree to finally give an estimate of execution time.

TO-DO :  
2. Describe the concept of multiple models for a single operator.
3. How do we select the best one among them. (Mention that current experiments don't use this idea, only one a simple non-linear model is used for all models with 2 features)

@@

%%%%%%%%%%%%%  This can be followed by several other sections
	\section{Training and Model selection}
	Before we describe the concept of multiple models for an operator, we shall present the motivation 
	for doing so. While the statistical models can find the complex non-linear dependencies between 			features and the target, they cannot model the interaction among operators which might
	affect the target value.

	Consider the classic example of Sorting, where in the target variable (execution time) is 
	proportional to $N_{l}$ log$N_{l}$. Unless this is used as a feature, the accuracy of the resulting 
	model can be disappointing. Since we don't know the right function to be used beforehand, 
	we explore the possible set of functions listed bellow and select the one with the 
	minimum estimation error. We've used functions that are similar to the one's defined in 					\cite{robustIISc}.

	\begin{itemize}
	\item Linear: $a_{0} N_{l} + a_{1} N_{r}$ 
	\item Quadratic: $a_{0} N_{l}N_{r}$ , $a_{0} N_{l}N_{l}$ , $a_{0} N_{r}N_{r}$ \ldots
	\item Logarithmic: $a_{0} N_{l} log(N_{r}) , a_{0} N_{r} log(N_{l}) $ .. 
	\item Exponential: $a_{0} N_{l} ^ {N_{r}}$  	
	\end{itemize}

	To illustrate the use of the functions in more details, we have plotted the curves with 
	different possible functions for QuickSort as well as IndexNestedLoop operator. For the purpose
	of generating curves for QuickSort we have used the following Query template

	\begin{framed}
	\begin{verbatim}
	Select * from orders
	where o_orderkey <= :varies 
	order by o_totalprice
	\end{verbatim}
	\end{framed}

	%We now show the observations for logarithmic and quadratic functions.
	\includegraphics[scale=0.38]{Plots/quicksortnlogn.png}	
	\includegraphics[scale=0.35]{Plots/quicksortnl2.png}

	Unsurprisingly, logarithmic function fits better than the quadratic. 
	While in some cases (Such as QuickSort) the appropriate function is obvious from SQL
	query processing, it does not hold for all operators. 
	For example, consider Nested Loop Join which 
	can be modeled with quadratic functions in general but when the inner relation has an index 
	the same model produces estimates estimates which are 
	"off" by orders of magnitude. In the following example,
	we take an exploratory approach to determine the best function for the Index Nested Loop operator.
	As in the case of QuickSort we have used the following query template to generate the required
	data.

	\begin{framed}
	\begin{verbatim}
	select * from lineitem,orders
	where l_orderkey = o_orderkey 
		  and l_orderkey<= : varies;
	\end{verbatim}
	\end{framed}

	\includegraphics[scale=0.37]{Plots/nil-nlnrlog2.png}	
	\includegraphics[scale=0.37]{Plots/nil-nl+nr.png}	
	\includegraphics[scale=0.37]{Plots/nil-nlnr.png}

	Out of the above 3 functions, $N_{l} log(N_{r})$ fits data better than other functions. 	
	\subsection{Best Fit Model}	
	Best fit model $M_{O}$ for an Operator $O$ is the model which has the minimum estimation errors 
	for the set of training queries. The estimation error is computed as follows,
	\\
	Model Estimation Error = \sum\limits_{i=1}^n \dfrac{\left| Actual_{i} - Estimated_{i} \right|}{n}

	\section{Preliminary Experiments}
	\subsection{Setup}
	Our experimental study uses TPC-H decision support benchmark \cite{TPCH} on the top of 
	PostgresSQL. The details are as follows,

	\begin{itemize}
	\item Database Management System : PostgreSQL 9.4
	\item Data sets and workload: We created 1GB TPC-H database
	according to the specification. The primary key indices as indicated
	in the TPC-H specification were created for both databases. We have excluded Query 1,21  
	To keep the training time under control, and Queries 15 as it creates a view which is not yet
	supported in our work. This resulted in 19 out of 22 TPC-H queries being used for training. We've 
	used TPC-H Qgen tool to generate 10 instances of each of the 19 queries, resulting in a total of 190 
	queries.

	\item Hardware: The queries were executed on a machine
	with 8GB RAM running Ubuntu. buffer pool size was set to 1GB (25\% of the total RAM as the rule
	of thumb). All queries were executed sequentially with cold start
	(i.e., OS buffers flushed before the start of each query).

	\item Statistical Models: As of now, we have limited our experiments to 2-features

	TODO : Finish tomo. Talk with sir once.

	TODO : 
	0 . Optimzer vs Actual. 
	1. Histogram of TPC-H errors.
	2. table with distribution of MRE.
	3. Performance operator wise. Box plot
	4. Insights : which operators are hard to predict

	\end{itemize}

	\subsection{Prediction with Optimizer cost models}

@@

	Give a background about Static and dynamic workload.  
	do experiments at 2 stages
	1. Act vs Act
	2. Est vs Est(Bias towards estimated cardinalities)
	For each type of above scenario, do testing for tpc-h queries 
	(tpc-ds in future)
	At different scales to see the model's ability to predict.
	For midterm you should be having result for 1 as well as 10 GB.
	For each above dataset, 
	Results should convey L1 (MRE) as well as queries that are within 
	[0-0.5] [0.5-1] [1-1.5]

%	Interpret tpc-h results by template wise, 
%	To provide a sense of the magnitude of the estimation errors made
%	by the optimizer, consider template-18, which is one of the templates 
%	that exhibit the biggest error in operator-level prediction with
%	actual/estimate model building. Instances of template-18 include 
%	the following group by clause on table lineitem:
%	group by l_orderkey having sum(l_quantity) > 314
%	There are 15 million distinct l_orderkey values in lineitem
%	(out of approximately 60 million tuples). The estimated number of
%	groups satisfying sum(l_quantity) > 314 is 399521, whereas
%	the actual number is 84. The PostgreSQL query optimizer com-
%	putes this estimate using histograms (with 100 bins) for each col-
%	umn based on the attribute independence assumption. The results
%	are later fed into a Hash-Semi-Join, whose cost estimate is correspondingly
%	very much off the mark.

	\section{Conclusions and Future Work}
	Talk about generalization. 
	operator specific features
	capturing query interaction i.e., pipeline source of over-estimation
	Optimizers provides sufficient information to detect a pipeline,
	 can we use it to bind the operators and predict execution time 
	 for a set of operators instead of a single operator.

	\begin{thebibliography}{9}
	\bibitem{ganapathi} 
	A. Ganapathi, H. A. Kuno, U. Dayal, J. L. Wiener, A. Fox, M. I. Jordan,
	and D. A. Patterson. Predicting multiple metrics for queries: Better
	decisions enabled by machine learning. In ICDE, 2009.

@@

	\bibitem{robustIISc}
	Harish, D., Pooja N. Darera, and Jayant R. Haritsa. "Identifying robust plans through plan diagram 			reduction." Proceedings of the VLDB Endowment 1.1 (2008): 1124-1140.

@@

		\bibitem{opthet} 
	Du, Weimin, Ravi Krishnamurthy, and Ming-Chien Shan. "Query optimization in a heterogeneous dbms." 		VLDB. Vol. 92. 1992.
	\bibitem{activeSLA}
	Xiong, Pengcheng, et al. "ActiveSLA: a profit-oriented admission control framework for database-as-a-		service providers." Proceedings of the 2nd ACM Symposium on Cloud Computing. ACM, 2011.	
	\bibitem{progress}
	Mishra, Chaitanya, and Nick Koudas. "The design of a query monitoring system." ACM Transactions on 		Database Systems (TODS) 34.1 (2009): 1.
